---
title: "BeadChip"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{beadchip}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(yamat)

library(minfi)

#' Get probe total number from IlluminaMethylationManifest object.
get_probe_total_from_manifest <- function(illumina_methylation_manifest) {
  probe_counts <- sapply(illumina_methylation_manifest@data, nrow)
  return(sum(probe_counts))
}


#' Get probe names from IlluminaMethylationManifest object.
get_probe_names_from_manifest <- function(illumina_methylation_manifest) {
  probe_names <- lapply(
    illumina_methylation_manifest@data,
    function(x) {x$Name}
  )
  return(do.call(c, probe_names))
}


#' Get probe counts from IlluminaMethylationManifest object.
get_probe_counts_from_manifest <- function(illumina_methylation_manifest) {
  probe_numbers <- sapply(illumina_methylation_manifest@data, nrow)
  unique_probe_names <- sapply(
    illumina_methylation_manifest@data, 
    function(x) {
      length(unique(x$Name))
    }
  )
  output <- data.frame(
    Probe_number = probe_numbers,
    Unique_probe_name = unique_probe_names
  )
  return(output)
}
```

# About

The document is about BeadChip and comparison.

# EPIC/850K

## Based on `IlluminaMethylationManifest`

```{r epic_manifest}
library(IlluminaHumanMethylationEPICmanifest)

IlluminaHumanMethylationEPICmanifest

# IlluminaMethylationManifest object
# Annotation
#   array: IlluminaHumanMethylationEPIC
# Number of type I probes: 142262 
# Number of type II probes: 724574 
# Number of control probes: 635 
# Number of SNP type I probes: 21 
# Number of SNP type II probes: 38 

str(IlluminaHumanMethylationEPICmanifest)

# Formal class 'IlluminaMethylationManifest' [package "minfi"] with 2 slots
#   ..@ data      :<environment: 0x000002336bb7b480> 
#   ..@ annotation: chr "IlluminaHumanMethylationEPIC"

get_probe_total_from_manifest(IlluminaHumanMethylationEPICmanifest)

# 867530

get_probe_counts_from_manifest(IlluminaHumanMethylationEPICmanifest)

#             Probe_number Unique_probe_name
# TypeII            724574            724574
# TypeSnpI              21                21
# TypeI             142262            142262
# TypeControl          635                 0
# TypeSnpII             38                38
```


```{r epic_anno}
library(tidyverse)
library(IlluminaHumanMethylationEPICanno.ilm10b2.hg19)
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b2.hg19)

dim(anno)
# [1] 866836     46

# Verify the code
anno %>%
  head() %>%
  as.data.frame() -> anno_test
rbind(anno_test, anno_test) %>%
  dplyr::filter(Type == "I") %>%
  dplyr::group_by(chr, pos) %>%
  dplyr::summarize(Count=n()) %>%
  dplyr::pull(Count) %>%
  table()

# 2 
# 6

# Six probes present twice.

anno %>%
  as.data.frame() %>%
  dplyr::filter(Type == "I") %>%
  dplyr::group_by(chr, pos) %>%
  dplyr::summarize(Count=n()) %>%
  dplyr::pull(Count) %>%
  table()

#      1 
# 142262

# All type I probes present once

anno %>%
  as.data.frame() %>%
  dplyr::filter(Type == "II") %>%
  dplyr::group_by(chr, pos) %>%
  dplyr::summarize(Count=n()) %>%
  dplyr::pull(Count) %>%
  table()

#      1 
# 724574

# All type II probes present once
```

Compare probe IDs/Names from manifest and annotation:

```{r}
probe_names_manifest <- get_probe_names_from_manifest(IlluminaHumanMethylationEPICmanifest)
probe_ids_annotation <- rownames(anno)

all(probe_names_manifest %in% probe_ids_annotation)

# [1] FALSE

all(probe_ids_annotation %in% probe_names_manifest)

# [1] TRUE

setdiff(probe_names_manifest, probe_ids_annotation)

#  [1] "rs10796216" "rs213028"   "rs951295"   "rs3936238"  "rs6991394" 
#  [6] "rs1520670"  "rs11034952" "rs9292570"  "rs2385226"  "rs1040870" 
# [11] "rs10882854" "rs5987737"  "rs13369115" "rs2208123"  "rs6471533" 
# [16] "rs715359"   "rs10033147" "rs5926356"  "rs6626309"  "rs654498"  
# [21] "rs10936224" "rs2468330"  "rs877309"   "rs2857639"  "rs798149"  
# [26] "rs939290"   "rs9839873"  "rs739259"   "rs133860"   "rs7746156" 
# [31] "rs1941955"  "rs1484127"  "rs2235751"  "rs6982811"  "rs2125573" 
# [36] "rs845016"   "rs4742386"  "rs1019916"  "rs9363764"  "rs472920"  
# [41] "rs1510189"  "rs6426327"  "rs1416770"  "rs2959823"  "rs2804694" 
# [46] "rs2032088"  "rs4331560"  "rs11249206" "rs1510480"  "rs264581"  
# [51] "rs6546473"  "rs2521373"  "rs10457834" "rs966367"   "rs7660805" 
# [56] "rs1495031"  "rs348937"   "rs1467387"  "rs1945975" 
```

All probe IDs in the annotation are presented in the manifest. The differences are SNP probes in the manifest but not the annotation.

```{r}
library(minfiDataEPIC)
dim(MsetEPIC)
# [1] 866836      3
```

## Based on Zhou's SeSAMe




# Comparisons

## EPIC v2 vs EPIC/850K

### Manifest from R packages

```{r}
library(tidyverse)
library(IlluminaHumanMethylationEPICanno.ilm10b2.hg19)
library(IlluminaHumanMethylationEPICv2anno.20a1.hg38)
anno_850 <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b2.hg19)
anno_epic_v2 <- getAnnotation(IlluminaHumanMethylationEPICv2anno.20a1.hg38)

# unique probe seqs
length(unique(c(anno_850$ProbeSeqA, anno_850$ProbeSeqB)))

# [1] 1009099


cg_number_epic_v2 <- sapply(anno_epic_v2$Name, function(x) { strsplit(x, split="_")[[1]][1] })
cg_number_epic_v2_850k_common <- intersect(anno_850$Name, cg_number_epic_v2)
length(cg_number_epic_v2_850k_common)

# 722272

anno_epic_v2 %>%
  as.data.frame() %>%
  dplyr::mutate(In_850K=stringr::str_starts(EPICv1_Loci , "cg")) %>%
  dplyr::group_by(In_850K) %>%
  dplyr::summarise(Count=n())

# A tibble: 2 × 2
#   In_850K  Count
#   <lgl>    <int>
# 1 FALSE   205691
# 2 TRUE    724384

anno_epic_v2 %>%
  as.data.frame() %>%
  dplyr::filter(!stringr::str_starts(EPICv1_Loci , "cg")) %>%
  dplyr::group_by(EPICv1_Loci) %>%
  dplyr::summarise(Count=n()) -> anno_epic_v2_new
```

### zwdzwd Infinium annotation

[Source](https://zwdzwd.github.io/InfiniumAnnotation)

```{r}
epic_v2_file <- "/Users/chenm8/Downloads/EPICv2.hg38.manifest.tsv.gz"
library(readr)
epic_v2_manifest <- readr::read_tsv(epic_v2_file)

table(epic_v2_manifest$CpG_chrm)
table(epic_v2_manifest$mapChrm_A)


# Unique position
epic_v2_manifest %>%
  as.data.frame() %>%
  dplyr::group_by(mapChrm_A , mapPos_A ) %>%
  dplyr::summarize(Count=n()) %>%
  dplyr::pull(Count) %>%
  table()
```



